\documentclass{bioinfo}
\usepackage{natbib}
\copyrightyear{2009}
\pubyear{2009}

\begin{document}
\firstpage{1}

\title[Modular analysis]{Modular analysis of gene expression data with
  GNU R or Matlab}
\author[G\'abor Cs\'ardi \textit{et~al}]{G\'abor Cs\'ardi\,,$^{1,2}$,
  Zolt\'an Kutalik\,$^{1,2}$ and Sven Bergmann\,$^{1,2}$}
\address{$^{1}$Department of Medical Genetics, and 
  $^{2}$Swiss Institute of Bioinformatics,
  University of Lausanne, Rue de Bugnon 27 - DGM 328, CH-1005 Lausanne,
  Switzerland.}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\maketitle

\begin{abstract}

\section{Summary:}
\section{Availability:}
\section{Contact:} \href{Sven.Bergmann@unil.ch}{Sven.Bergmann@unil.ch}
\end{abstract}

\section{Introduction}

What is biclustering.
What is the ISA.

\section{Methods}

Normalization. 

Random and smart seeding. 

The ISA iteration. 

Merging the modules.

The robustness measure and filtering the modules. 

Module trees.

\section{R Implementation}

Two packages: isa2 does the computation, eisa does everything related
to biology.

\subsection{Workflow}

Two workflows, the simple workflow does everything with default
parameters. In the detailed workflow the user has access to every step
of the analysis.

\subsection{Visualization}

Expression plots, GO tree plots, module tree plots, HTML summary
generation.

\subsection{Connection to other software}

The biclust package.

\section{Matlab implementation}

\section*{Acknowledgement}

\paragraph{Funding\textcolon} 

\paragraph{Conflict of interest\textcolon} None declared.

\bibliographystyle{natbib}
\bibliography{isa}
\nocite{*}

\end{document}
